---
title: "Activity Monitoring data Analysis"
date: "Saturday, December 13, 2014"
output: html_document
---

```{r,echo=FALSE}

library(lattice)
library(ggplot2)
```
####Loading and Preprocessing the data 

Load data from the current working directory.  

```{r filereadstore}
 monitor_data <- read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
monitor_data$date <- as.Date(monitor_data$date, "%Y-%m-%d")
head(monitor_data,5)
```

####Make a histogram of the total number of steps taken each day

```{r makrhist}
 Totsteps <- aggregate(steps ~ date, data = monitor_data, sum, na.rm = TRUE)
hist(Totsteps$steps, main = "Total steps in Days", xlab = "Day", col = "blue")
```

####mean and median total number of steps taken per day
```{r mean&edian}
 myFun <- function(x){ 
   c(mean=mean(x), median = median(x))}
myFun(Totsteps$steps)
```

####What is the average daily activity pattern?
#####Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r dailyact,fig.height=6}
average_daily <- aggregate(monitor_data$steps,  by = list(interval = monitor_data$interval), FUN=mean, na.rm=TRUE)
colnames(average_daily) <- c("interval", "steps")
x <- average_daily$interval
y <- average_daily$steps
plot(x, y, type = "l", xlab = "Interval in min.", 
     ylab = "Average no of steps", main = "Average Activity pattern", 
     col = "red")
```
#####5-minute interval having the maximum number of steps
```{r maxaverage}
max_average <- average_daily[which.max(average_daily$steps),]$interval
```

####Imputing missing values
##### Total number of missing values in the data set
```{r missingvalues}
sum(is.na(monitor_data))
```
#####Input the missing values with average 5 min. interval
```{r addmissingvalue}
data_na <- monitor_data 
for(i in 1:nrow(data_na))
  {if(is.na(data_na$steps[i])){
    data_na$steps[i] <- average_daily[which(data_na$interval[i]==    average_daily$interval),]$steps
  }
  
  }


```
#####Histogram with the new dataset with replace na values
```{r }
 Totsteps_new <- aggregate(steps ~ date, data = data_na, sum, na.rm = TRUE)
hist(Totsteps_new$steps, main = "Total steps in Days", xlab = "Day", col = "blue")
```
#####new mean and median values.
######The missing values has minimal impact on mean median
```{r }
myFun(Totsteps_new$steps)
```
####Activity patterns between weekdays and weekends
#####Create the data set indicating weekdays& weekends
```{r,fig.height=4}

day_name <- weekdays(data_na$date)
daylevel <- vector()
for (i in 1:nrow(data_na)) {
    if (day_name[i] == "Saturday" || day_name[i]== "Sunday") 
        daylevel[i] <- "Weekend"
     else 
        daylevel[i] <- "Weekday"
    
}
data_na$daylevel <- daylevel
data_na$daylevel <- factor(data_na$daylevel)
average_daily_new <- aggregate(data_na$steps,  by = list(interval = data_na$interval,daylevel = data_na$daylevel), FUN=mean, na.rm=TRUE)
colnames(average_daily_new) <- c("interval","daylevel", "steps")
xyplot(steps~ interval|daylevel,average_daily_new,type="l",layout=c(1,2),xlab ="Interval",ylab ="Number of Steps",col="red")



```
